The truncated KdV framework for modeling anomalous waves induced by abrupt depth changes

### 2019-08-08 paper2_JNLS/paper.tex
NM

Rogue waves are abnormally large surface waves, defined by oceanographers as those that exceed twice the significant wave height. Though such waves were once dismissed as myth, they have now been recorded in oceans across the globe and pose a recognized threat to seagoing vessels and naval structures. Rogue waves, also variously known as freak or anomalous waves, have been observed both in deep and shallow water, and can be triggered by a variety of mechanisms, including anomalous wind-forcing, focusing due to variable bathymetry, and a deep-water modulational instability that goes by the name Benjamin-Feir. The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement. That is, when governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these rare events can become much more common when surface statistics deviate from Gaussian.

@@

Our particular focus is the laboratory experiments of Bolles et al.~\cite{bolles2019anomalous}, who demonstrated the emergence of anomalous statistics in a randomized wave-field that crosses an abrupt depth change (ADC). In these experiments, an incoming wave-field was generated with nearly Gaussian statistics. A plexiglass step created the abrupt depth change. Upon passing into a region of shallower depth, the wave distribution skewed strongly, with deviation from Gaussian being most pronounced a short distance downstream of the ADC. 
Inspired by these experimental observations, Majda et al.~\cite{majda2019statistical} laid out a theoretical foundation that accurately captures several aspects of the anomalous behavior. The theory is based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) equations, and uses a combination of computational, statistical, and analytical tools. In follow-up work, Majda and Qi (cite) performed subsequent analysis on severely truncated, low-dimensional systems that were shown to exhibit many of the same anomalous features.

The purpose of the present manuscript is to expand on both the experimental findings of Bolles et al.~\cite{bolles2019anomalous} and the theoretical analysis of Majda et al.~\cite{majda2019statistical}, with particular emphasis on the synergy between the two; e.g.~model developments inspired by experimental findings, and new experimental tests inspired by theoretical predictions. We present a range of previously unpublished experimental data, including statistics of the surface {\em slopes} (as opposed to only displacement data as published in \cite{bolles2019anomalous}, autocorrelation data, and some data on experiments in which the depth ratio was varied. We also expand on the theoretical treatment of \cite{majda2019statistical}, with emphasis on more direct link to experiments.

>>>>

Rogue waves are abnormally large surface waves, defined by oceanographers as those that exceed twice the significant wave height. Though such waves were once dismissed as myth, they have now been recorded in oceans across the globe and pose a recognized threat to seagoing vessels and naval structures. Rogue waves, also variously known as freak or anomalous waves, have been observed both in deep and shallow water, and can be triggered by a variety of mechanisms, including anomalous wind-forcing, focusing due to variable bathymetry, and a deep-water modulational instability that goes by the name Benjamin-Feir. The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement; when governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these rare events can become much more common when surface statistics deviate from Gaussian.

@@

Our particular focus is the laboratory experiments of Bolles et al.~\cite{bolles2019anomalous}, who demonstrated the emergence of anomalous statistics in a randomized wave-field upon encountering an abrupt depth change (ADC). In these experiments, an incoming wave-field is generated with nearly Gaussian statistics, and a plexiglass step placed near the middle of the tank creates the depth change. Upon passing into the region of shallower depth, the wave distribution skews strongly, with deviation from Gaussian being most pronounced a short distance downstream of the ADC. 
Inspired by these experimental observations, Majda et al.~\cite{majda2019statistical} laid out a theoretical foundation that accurately captures several aspects of the anomalous behavior. The theory is based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) equations, and uses a combination of computational, statistical, and analytical tools. In follow-up work, Majda and Qi (cite) performed analysis on more severely truncated, lower-dimensional systems that were shown to exhibit many of the same anomalous features.

The purpose of the present manuscript is to expand on both the experimental findings of Bolles et al.~\cite{bolles2019anomalous} and the theoretical analysis of Majda et al.~\cite{majda2019statistical}, with particular emphasis on the synergy between the two; e.g.~experimental discoveries guiding theoretical developments and theoretical predictions inspiring new experimental tests. We present a range of previously unpublished experimental data, including statistics of the surface {\em slopes} (as opposed to only displacement data as published in \cite{bolles2019anomalous}, autocorrelation data, and some data on experiments in which the depth ratio was varied. We also expand on the theoretical treatment of \cite{majda2019statistical}, with emphasis on more direct link to experiments.
